\section{New Cutoff Results} \label{gua:sec:cutoffs}

We present new cutoff results that extend Theorems~\ref{thm:disj-cutoff-pairs} and \ref{thm:conj-cutoff}.
The new and previous results are summarized in the table below.

\begin{table}[h]
\centering
%\vspace{-10pt}
\label{table:cutoffs}
\centering
\setlength{\tabcolsep}{2pt}
{%\resizebox{\linewidth}{!}{
\begin{tabular}{ r|c|c|c|c }
   & \specialcellC{$h(A,B^{(k)})$ \\ no fairness} & \specialcellC{deadlock detection \\ no fairness} & \specialcellC{$h(A,B^{(k)})$ \\ uncond. fairness} & \specialcellC{deadlock detection \\ strong fairness}  \\[9pt]
\hline
Disjunctive~ & $|B| + k + 1$ &
        $2|B| - 1$ & 
        $2|B| + k - 1$ &
        $2|B| - 1$ 
          \\[4pt]
\hline
Conjunctive~ & 
        $k+1$ &  
        $2|B|-2~(*)$ & 
        $k+1~(*)$ &  
        $2|B|-2~(*)$
\end{tabular}
}
%\vspace{-10pt}
\end{table}
The table distinguishes between disjunctive and conjunctive systems (in rows).
In the columns,
we consider satisfaction of properties $h(A,B^{(k)})$ and the existence of deadlocks,
with and without fairness assumptions.
All results hold for open systems, and for both path quantifiers $\pforall$ and $\pexists$.
Cutoffs depend on the size of process template $B$ and the number $k \geq 1$ of $B$-processes a property talks about.

Results marked with a $(*)$ are for a restricted class of systems:
for conjunctive systems with fairness, we require infinite runs to be 
initializing, i.e., all non-deadlocked 
%$B$-processes 
processes
return to 
$\init$
%$\initstate_B$ 
infinitely often.\footnote{This assumption is in the same 
flavor as the restriction that $\initstate_A$ and $\initstate_B$ appear in 
all conjunctive guards. Intuitively, the additional restriction makes sense 
since conjunctive systems model shared resources, and everybody who takes a 
resource should eventually release it.}
Additionally, the cutoffs for deadlock detection in conjunctive systems only support $1$-conjunctive systems.
\ak{explain why we need it}
%The reason for this restriction will be explained in Remark~\ref{rem:general-conj-tough} and Appendix~\ref{gua:sec:app-conj}.

All cutoffs in the table are tight---no smaller cutoff can exist for 
this class of systems and properties---except for the case of deadlock 
detection in disjunctive systems without fairness. There, the cutoff is 
asymptotically tight, i.e., it must increase linearly with the size 
of the process template.

Note that the table does not describe all possible combinations:
for example, we do not consider satisfaction of $h(A,B^{(k)})$ on strong-fair runs.
But the results in the table are the most interesting, from our view,
for parameterized synthesis.

In the following sections we prove the results.
