\section{Related Work} \label{gua:sec:related}
\ak{ cite ``Parameterized Model-Checking of Timed Systems with Conjunctive Guards'' and another Sasha's paper that contain results for disj guards (including undec result for some CTL*-like logic). }

In this work we extend the results of Emerson and Kahlon~\cite{Emerson00} who study PMC of guarded protocols, but do not support fairness assumptions, nor provide cutoffs for 
deadlock detection.
In \cite{EmersonK03} they extended their work to systems with limited forms of guards and broadcasts, and also proved undecidability 
of PMC of conjunctive guarded protocols wrt.\ $\LTL$ (including $\nextt$), 
and undecidability wrt.\ $\LTLmX$ for systems with both 
conjunctive and disjunctive guards.

Bouajjani et al.~\cite{Bouajjani08} study parameterized model checking of
resource allocation systems (RASs). Such systems have a bounded number of resources, 
each owned by at most one process at any time. Processes are pushdown automata, 
and can request resources with high or normal priority.
RASs are similar to conjunctive guarded protocols in that certain
transitions are disabled unless a processes has a certain resource. 
RASs without priorities and where all the processes are finite state Moore machines can 
be converted to conjunctive guarded protocols (at the price of blow up), 
but not vice versa.
The authors study parameterized model checking wrt.\ $\LTLmX$ properties under certain fairness assumptions,
and deadlock detection.
%The authors study parameterized model checking wrt.\ $\LTLmX$ properties on arbitrary or on strong-fair runs,
%and (local or global) deadlock detection.
Their proofs are based on ideas of~\cite{Emerson00} (our proofs are also based on ideas of~\cite{Emerson00}).
 
German and Sistla~\cite{German92} considered global deadlocks and strong fairness properties for systems with pairwise rendezvous communication in a clique.
In such systems, processes communicate pairwise using messages:
one process sends a message and blocks until another process reads the message.
Emerson and Kahlon~\protect\cite{EmersonK03} have shown that disjunctive guard systems can be reduced to such pairwise rendezvous systems.
However, German and Sistla \cite{German92} do not provide cutoffs, 
nor do they consider deadlocks for individual processes,
and their specifications can talk about one process only.
%
Aminof et al.~\cite{AminofKRSV14} have 
recently extended these results to more general topologies, and have 
shown that for some decidable parameterized model checking problems there are no cutoffs,
even in cliques.
\ak{update the paragraphs, they also have some results on disj?}

Many of the decidability results above have been surveyed in our book~\cite{BloemETAL15}.
