\part[Excursion into Parameterized Systems]{Excursion Into\\Parameterized Systems\\ \ \\ \LARGE Guarded and Token-ring Systems}

\section*{Overview of Part II}

Concurrent systems are hard to implement and even harder to debug.
On the other side, they are relatively easy to specify.
Consider, for example, the arbiter serving many clients.
A possible specification is
$
  \forall i \neq j. \G \neg ( g_i \land g_j ) \land \G (r_i \impl \F g_i),
$
%\[
%\begin{array}{ll}
%  \forall i \neq j.~ & \G \neg ( g_i \land g_j ) \land \\
%  \forall i.~ & \G (r_i \impl \F g_i),
%  \end{array}
%\]
which says that, for every client, every request should be eventually granted,
and the grants are mutually exclusive.
If a human implements such an arbiter,
he would try to come up with a basic block that handles a single client,
and connect such a block into a system, that handles as many clients as needed.
On the other side,
the computer tries to synthesize a system as one monolithic block.
This hides the insight that
a system for $n+1$ clients is very similar to a system for $n$ clients.
This leads to the scalability problem, once we require a large number of clients.

The parameterized synthesis approach~\cite{JB14} addresses the issue.
The idea is---just like the human would do---%
to automatically synthesize a basic block that
can be arranged into a system of any desired size.
There are several ways to arrange such blocks into a system,
depending on how they communicate with each other.
In this thesis part we will look into two system architectures.

The first architecture is inspired by cache coherence protocols found in modern processors.
Such a protocol is described by states,
where transitions between states happen depending on whether or not there is a processor in a particular state.
I.e., the transitions are guarded.
Chapter~\ref{chap:guarded-systems} studies guarded systems.

The second kind of systems is token-ring systems.
In such a system, the single token circulates in the system.
A process possessing the token knows that no other process has the token.
Based on this information, the process can, for example, raise the grant.
If all processes raise the grant only when they posses the token,
then the grants will be mutually exclusive.
Chapter~\ref{chap:token-systems} studies token-ring systems.

For both architectures we study their parameterized synthesis problems.
The parameterized synthesis problems asks, given a parameterized specification,
to find a process implementation,
such that a system of any size composed of such processes,
satisfies the specification.
The solution to the seemingly difficult problem---we now ask for correctness of a system of \emph{any} size---is based on the cutoff reduction:
to synthesize a process that works for all system sizes,
it is enough to synthesize a process that works in a system of a cutoff size.
For example, for the specification of the arbiter mentioned above,
the cutoff for token-ring systems is 4.
This means that it is enough to find a process implementation that
works in a system with 4 such processes.
Once we find it, a system of size 5, 6, 7,... is also correct.

In Chapter~\ref{chap:guarded-systems} we prove cutoff results for guarded systems.
Our results extend the results of Emerson and Kahlon~\cite{EmersonK03}.
Our contribution concerns both parameterized synthesis and parameterized verification.
We prove new cutoff results that are applicable to a previously unconsidered setting
of open systems with liveness properties under fairness assumptions.
We also prove new cutoff results for deadlock detection.
The work is theoretical; it is yet to find its application.

In Chapter~\ref{chap:token-systems} on token-ring systems,
we extend the cutoffs of Emerson and Namjoshi~\cite{Emerso03}
to a new setting of fully asynchronous systems and richer specifications.
Then we apply them to an industrial arbiter protocol called AMBA.
Thus, we synthesize for the first time the AMBA protocol in the parameterized sense.

The chapters can be read in any order.

\input{guarded-systems/main}

\input{token-systems/main}
